---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

# Working on

### *1. Building in statistic framework*

------------------------------------------------------------------------

# Initialization and Import

Define experimental parameters and input/output information

```{r Datapath, output, and modules}
# Define raw data directory
data_path = 'csvs/'
# Define experiment output name
output_name <- 'Sox10MO_smpd3Intron'

# Import modules
library(dplyr)            # needed for %>%
library(tidyverse)
library(tidyr)             # contains "separate" function
library(ggplot2)
library(GGally)
# install.packages("ggbeeswarm")
library(ggbeeswarm)

```

Navigate to raw data from Fiji and import the raw measurement files

```{r Import raw files}
# Loop through files in source data to pull in Fiji measurements
file_list = list.files(path = data_path)     # Determine files to loop through

for (file_ in file_list) {
  if (file_ == file_list[1]){   # Pull in first file to initialize df
    df_raw <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    df_raw['File'] <- file_
  }
  else {                       # Pull in the rest of the files and concatenate
    temp_data <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    temp_data['File'] <- file_
    df_raw <- rbind(df_raw, temp_data)  # Concatenates temp_data with df
  }
}
head(df_raw)
```

Parse out the relevant metadata from the file name and Label columns, then discard unnecessary columns

```{r Parse metadata from file and label information}
df <- df_raw %>% separate(Label, c('Channel', 'ROI'), ":")
df <- df %>% separate(File, c('Date','Treatment',
                              'Stains','Embryo','Somites', 'Section'), "_")
df <- df %>% separate(Section, c('Section'))  # This deletes the .csv file extension
df <- unite(df, Emb_ID, c('Date','Treatment','Embryo'), sep = '_', remove = FALSE)
df <- unite(df, Sec_ID, c('Date','Treatment','Embryo','Section'), sep = '_', remove = FALSE)
df <- select(df, -c('Stains')) # Drop unnecessary cols
head(df)
```

# Normalize and Analyze

## SMPD3 Intron HCR Integrated Density Fluorescence Analysis

Analysis steps:

1.  Subset df to display only SMPD3 intron HCR values for control and experimental ROIs only

2.  Generate a ratio per section and prepare a df for "per section" plots

3.  Average section ratios and make a new df for "per embryo" plots

```{r Subset and tidy data for "per section" values}
# Get rid of unnecessary ROIs and Channels
data_section_intden <- df[!(df$ROI=="Background"),]
data_section_intden <- data_section_intden[(data_section_intden$Channel == "smpd3_intron"),]
data_section_intden <- select(data_section_intden, -c('Area', 'Mean', 'RawIntDen')) # Drop unnecessary cols

# Pivot table
data_section_intden <- data_section_intden %>% pivot_wider(names_from = 'ROI', values_from = 'IntDen')

# Determine the ratio of Experimental/Control IntDen
data_section_intden['Ratio'] <- data_section_intden['ExptArea'] /     data_section_intden['CntlArea']

print(mean(data_section_intden$Ratio))
head(data_section_intden)
```

```{r Subset and tidy data for "per embryo" values}
data_embryo_intden <- data_section_intden
data_embryo_intden <- data_embryo_intden %>% group_by(Emb_ID) %>% summarize(Cntl = mean(CntlArea), Expt = mean(ExptArea), Ratio = mean(Ratio))

head(data_embryo_means)
```

Save out resulting data

```{r}
date <- Sys.Date()
date <- format(date, format="%Y%m%d")
write.csv(data_section_intden, file=paste(date, output_name, 'IntDen_perSection.csv', sep = "_", collapse = NULL))
write.csv(data_embryo_intden, file=paste(date, output_name, 'IntDen_perEmbryo.csv', sep = "_", collapse = NULL))
```

## SMPD3 Intron HCR CTCF Analysis

Analysis steps:

1.  Subset df to display only SMPD3 intron HCR values

2.  Calculate CTCF values:

    CTCF = ROI IntDen - (ROI Area \* Background Mean)

3.  Generate a ratio per section and prepare a df for "per section" plots

4.  Average section ratios and make a new df for "per embryo" plots

```{r Subset and tidy data for "per section" values}
# Get rid of unnecessary ROIs and Channels
data_section_ctcf <- df[(df$Channel == "smpd3_intron"),]
# backgrounds <- data_section_ctcf[(data_section_ctcf$ROI =='Background'),]
# data_section_ctcf <- data_section_ctcf[!(data_section_ctcf$ROI=="Background"),]

data_section_ctcf <- select(data_section_ctcf, -c('RawIntDen')) # Drop unnecessary cols

# Pivot table
data_section_ctcf <- data_section_ctcf %>% pivot_wider(names_from = 'ROI', values_from = c('IntDen', 'Area', 'Mean'))

data_section_ctcf['CntlCTCF'] <- data_section_ctcf['IntDen_CntlArea'] -  (data_section_ctcf['Area_CntlArea'] * data_section_ctcf['Mean_Background'])

data_section_ctcf['ExptCTCF'] <- data_section_ctcf['IntDen_ExptArea'] - (data_section_ctcf['Area_ExptArea'] * data_section_ctcf['Mean_Background'])

# Determine the ratio of Experimental/Control IntDen
data_section_ctcf['Ratio'] <- data_section_ctcf['ExptCTCF'] /data_section_ctcf['CntlCTCF']

print(mean(data_section_ctcf$Ratio))
head(data_section_ctcf)
```

```{r Subset and tidy data for "per embryo" values}
data_embryo_ctcf <- data_section_ctcf
data_embryo_ctcf <- data_embryo_ctcf %>% group_by(Emb_ID) %>% summarize(Cntl = mean(CntlCTCF), Expt = mean(ExptCTCF), Ratio = mean(Ratio))

head(data_embryo_ctcf)
```

Save out resulting data

```{r}
date <- Sys.Date()
date <- format(date, format="%Y%m%d")
write.csv(data_section_ctcf, file=paste(date, output_name, 'CTCF_perSection.csv', sep = "_", collapse = NULL))
write.csv(data_embryo_ctcf, file=paste(date, output_name, 'CTCF_perEmbryo.csv', sep = "_", collapse = NULL))
```
