---
title: "Enh3 Intensity Analysis"
input: .csv files with measured ROI intensity from Fiji
output: .csv file with normalized values
---

Define parameters and import modules for this analysis 
```{r}
# Necessary input parameters
data_path = 'raw_source_data/'
measure_channel <- 'Enh3GFP'
norm_channel <- 'H2BRFP'
output_name <- 'Sox9MO_Enh3Intensity_results.csv'

# Import modules
library(dplyr)
library(tidyverse)
library(tidyr)
```

Navigate to raw data from Fiji and import the raw measurment files
```{r}
# Loop through files in source data to pull in Fiji measurements
file_list = list.files(path = data_path)     # Determine files to loop through

for (file_ in file_list) {
  if (file_ == file_list[1]){   # Pull in first file to initialize df
    df <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    df['File'] <- file_
  }
  else {                       # Pull in the rest of the files and concatenate
    temp_data <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    temp_data['File'] <- file_
    df <- rbind(df, temp_data)  # Concatenates temp_data with df
  }
}
head(df,5)
```

Parse out the relevant metadata from the file name and Label columns, then discard unnecessary columns
```{r}
df <- df %>% separate(Label, c('Channel', 'ROI'), ":")
df <- df %>% separate(File, c('Date','Treatment',
                              'Stains','Embryo','Somites','Mag','Type'), "_")
df <- unite(df, Emb_ID, c('Date','Treatment','Embryo'), sep = '_')
df <- select(df, -c('Min','Max','RawIntDen','Stains','Mag','Type')) # Drop unnecessary cols

```

Normalization and analysis pipeline:
  1: Channel of interest / Electroporation normalization channel
  2: Explicit value / mean of the control dataset (normalizes the control dataset around 1)
  3: Generate ratio of Experimental / Control values
```{r}
# Parse out and normalize data
data <- df[df$Channel==measure_channel & df$ROI!='Background',]
data['NormIntDen'] <- (data['IntDen'] / df[df$Channel==norm_channel 
                                    & df$ROI!='Background',]['IntDen'])
data <- select(data, -c('Area', 'Mean', 'IntDen')) # Drop unnecessary cols
data <- data %>% pivot_wider(names_from='ROI', values_from='NormIntDen')
data <- data %>% dplyr::rename('Cntl'='CntlArea', 'Expt'='ExptArea')

# Compute mean of control dataset and normalize everything to that
control_mean <- mean(data$Cntl)
data$Cntl <- data$Cntl/control_mean
data$Expt <- data$Expt/control_mean
# Determine the ratio of Experimental/Control IntDen
data$Ratio <- data$Expt/data$Cntl
head(data,5)
```

Save out resulting data
```{r}
write.csv(data, file=paste(Sys.Date(), output_name, sep = "_", collapse = NULL))
```
