---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

# Working on

### *1. Make loops to normalize within given treatments, doses, and/or stages*

### *2. Building in statistic framework*

3.  Practice using R to analyze migration area

------------------------------------------------------------------------

# Initialization and Import

Define experimental parameters and input/output information

```{r}
# Define raw data directory
data_path = 'csvs/'
# Define experiment output name
output_name <- 'ApoDMO_DR_MigrationArea_results.csv'

# Import modules
library(dplyr)            # needed for %>%
library(tidyverse)
library(tidyr)             # contains "separate" function
library(ggplot2)
library(GGally)
# install.packages("ggbeeswarm")
library(ggbeeswarm)

```

Navigate to raw data from Fiji and import the raw measurement files

```{r}
# Loop through files in source data to pull in Fiji measurements
file_list = list.files(path = data_path)     # Determine files to loop through

for (file_ in file_list) {
  if (file_ == file_list[1]){   # Pull in first file to initialize df
    df_raw <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    df_raw['File'] <- file_
  }
  else {                       # Pull in the rest of the files and concatenate
    temp_data <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    temp_data['File'] <- file_
    df_raw <- rbind(df_raw, temp_data)  # Concatenates temp_data with df
  }
}
head(df_raw)
```

Parse out the relevant metadata from the file name and Label columns, then discard unnecessary columns

```{r}
df <- df_raw %>% separate(Label, c('Channel', 'ROI'), ":")
df <- df %>% separate(File, c('Date','Treatment',
                              'Stains','Embryo','Somites', 'Mag'), "_")
df <- df %>% separate(Treatment, c('Treatment', 'Dose'), ";")

df <- unite(df, Emb_ID, c('Date','Treatment','Dose','Embryo'), sep = '_', remove = FALSE)
df <- select(df, -c('Channel','Mean', 'IntDen', 'RawIntDen',
                    'Stains','Mag')) # Drop unnecessary cols
head(df)
```

# Normalize and Analyze

Analysis steps:

1.  Pivot to position Control and Experimental areas adjacent one another

2.  Normalize to control dataset = Explicit value / mean of the control dataset (defines the mean of the control dataset as 1) - USED FOR PARALLEL COORDINATE PLOTS

3.  Generate ratio of Experimental / Control values - USED FOR BOX AND WHISKER/JITTER PLOTS

```{r}
# Get rid of unnecessary ROIs and Channels
data <- df[!(df$ROI=="Background"),]

# Pivot table
data <- data %>% pivot_wider(names_from = 'ROI', values_from = 'Area')

# Loop through conditions (date, dose, treatment, etc)
# First get a list of conditions to be normalized separate from one another
dose_list <- unique(data$Dose)

# Initialize new output DF and loop through each condition
data_final_list = list()
for (dose_ in dose_list) {
  data_temp <- data[(data$Dose==dose_),]
  # Normalize each value to the mean of the Control dataset
  data_temp['NormCntlArea'] <- data_temp['CntlArea'] / mean(data_temp$CntlArea)
  data_temp['NormExptArea'] <- data_temp['ExptArea'] / mean(data_temp$CntlArea)
  data_final_list <- append(data_final_list, list(data_temp))
}
data_final <- data_final_list %>% reduce(bind_rows)

# Determine the ratio of Experimental/Control IntDen
data_final['Ratio'] <- data_final['ExptArea'] / data_final['CntlArea']

head(data)
```

Save out resulting data

```{r}
date <- Sys.Date()
date <- format(date, format="%Y%m%d")
write.csv(data_final, file=paste(date, output_name, sep = "_", collapse = NULL))
```

# Plot Data

#### Parallel Coordinate Plots:

Good for experiments with paired data (i.e. bilateral electroporations)

```{r}
# Parallel coordinate plot
p <-ggparcoord(data_final, columns= 9:10,  showPoints=TRUE, 
      scale='globalminmax',
      groupColumn = 4,
      ) +
      labs(x ='Treatment', 
          y = 'Migration Area \n normalized to control') +
      scale_x_discrete(labels= c("NormCntlArea"="Control MO",
                                 "NormExptArea"="ApoD MO")) +
      ylim(0,2) +
      theme_classic() +
      theme(axis.text = element_text(size=10, color = "black")) 
      # theme(legend.position = "none")
p
```

#### Box and whisker/Jitter Plot

Good for experiments with categorical data or comparing ratios

```{r}
# Box and whiskers with strip plot for ratios
q <-ggplot(data_final, aes(x=Dose, y=Ratio)) +
      geom_boxplot() +
      geom_beeswarm(size=2, cex = 4, 
              color=data_final$Somites, show.legend = TRUE) +
      labs(x ='Treatment', 
          y = 'Norm. Migration Area') +
      ylim(0,1.75)  +
      theme_classic() +
      geom_hline(yintercept=1, linetype="dashed", color = "red") +
      theme(axis.text = element_text(size=10, color = "black")) + 
      theme(legend.position = "right")
     
q
```
