---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

# Working on

### *1. Make loops to normalize within given treatments, doses, and/or stages*

### *2. Building in statistic framework*

3.  Practice using R to analyze migration area

------------------------------------------------------------------------

# Initialization and Import

Define experimental parameters and input/output information

```{r}
# Define raw data directory
data_path = 'test_data/'
# Define experiment output name
output_name <- 'Sox9MO_MigrationArea_results.csv'

# Import modules
library(dplyr)            # needed for %>%
library(tidyverse)
library(tidyr)             # contains "separate" function
library(ggplot2)
library(GGally)
# install.packages("ggbeeswarm")
library(ggbeeswarm)

```

Navigate to raw data from Fiji and import the raw measurement files

```{r}
# Loop through files in source data to pull in Fiji measurements
file_list = list.files(path = data_path)     # Determine files to loop through

for (file_ in file_list) {
  if (file_ == file_list[1]){   # Pull in first file to initialize df
    df_raw <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    df_raw['File'] <- file_
  }
  else {                       # Pull in the rest of the files and concatenate
    temp_data <- read.csv(file.path(data_path, file_), header=TRUE, row.names=1)
    temp_data['File'] <- file_
    df_raw <- rbind(df_raw, temp_data)  # Concatenates temp_data with df
  }
}
head(df_raw)
```

Parse out the relevant metadata from the file name and Label columns, then discard unnecessary columns

```{r}
df <- df_raw %>% separate(Label, c('Channel', 'ROI'), ":")
df <- df %>% separate(File, c('Date','Treatment',
                              'Stains','Embryo','Somites','Mag','Type'), "_")
#df <- df %>% separate(Treatment, c('Gene', 'Dose'), ";")

df <- unite(df, Emb_ID, c('Date','Treatment','Embryo'), sep = '_', remove = FALSE)
df <- select(df, -c('Mean','Min','Max', 'IntDen', 'RawIntDen',
                    'Stains','Mag','Type')) # Drop unnecessary cols
head(df)
```

# Normalize and Analyze

Analysis steps:

1.  Pivot to position Control and Experimental areas adjacent one another

2.  Normalize to control dataset = Explicit value / mean of the control dataset (defines the mean of the control dataset as 1) - USED FOR PARALLEL COORDINATE PLOTS

3.  Generate ratio of Experimental / Control values - USED FOR BOX AND WHISKER/JITTER PLOTS

```{r}
# Get rid of unnecessary ROIs and Channels
data <- df[!(df$ROI=="Background"),]
data <- data[(data$Channel=="Pax7"),]

# Pivot table
data <- data %>% pivot_wider(names_from = 'ROI', values_from = 'Area')

# Normalize each value to the mean of the Control dataset
data['NormCntlArea'] <- data['CntlArea'] / mean(data$CntlArea)
data['NormExptArea'] <- data['ExptArea'] / mean(data$CntlArea)

# Determine the ratio of Experimental/Control IntDen
data['Ratio'] <- data['ExptArea'] / data['CntlArea']

head(data)
```

Save out resulting data

```{r}
date <- Sys.Date()
date <- format(date, format="%Y%m%d")
write.csv(data, file=paste(date, output_name, sep = "_", collapse = NULL))
```

# Plot Data

#### Parallel Coordinate Plots:

Good for experiments with paired data (i.e. bilateral electroporations)

```{r}
# Parallel coordinate plot
p <-ggparcoord(data, columns= 9:10,  showPoints=TRUE, 
      scale='globalminmax',
      groupColumn = 4,
      ) +
      labs(x ='Treatment', 
          y = 'Migration Area \n normalized to control') +
      scale_x_discrete(labels= c("NormCntlArea"="Control MO",
                                 "NormExptArea"="Sox9 MO")) +
      ylim(0,2) +
      theme_classic() +
      theme(axis.text = element_text(size=10, color = "black")) +
      theme(legend.position = "none")
p
```

#### Box and whisker/Jitter Plot

Good for experiments with categorical data or comparing ratios

```{r}
# Box and whiskers with strip plot for ratios
q <-ggplot(data, aes(x=Treatment, y=Ratio)) +
      geom_boxplot() +
      geom_beeswarm(size=4, cex = 4) +
      labs(x ='Treatment', 
          y = 'Norm. Migration Area') +
      scale_x_discrete(labels= c("Sox9MO05mM"="Sox9 MO")) +
      ylim(0,1.25)  +
      geom_hline(yintercept=1, linetype="dashed", color = "red") +
      theme_classic() +
      theme(axis.text = element_text(size=10, color = "black"))
    
q
```
